
#Script to analyse the frequency data and associate with features for Zanini data. 

* Read the csv files 
* Perform necessary calcuations
* Plot results (eventually new script)

Make sure to set your working directory to the correct folder
Kristof: setwd('/Users/hopla/project-master/project/gitpleuni/bachelerProject/Rscripts')
PSP Oct 2016: changed setwd command to go to Dropbox shared folder. 


```{r}
#setwd("~/Documents/Git/bachelerProject/Rscripts")
setwd("~/Dropbox/2016Project_Gabriella _Abdul/Rscripts")
source('./baseRscript.R')
library(scales)
library(plotrix)

#Gabriella: setwd("C:/Users/Gaby/Dropbox/2016Project_Gabriella _Abdul/Rscripts")
#10/12/16 might need to install package Pegas in order to run other programs in baseRscript.R

```

Read the stored frequencies rather than calculating frequencies again
```{r}
read.table("../Output/freqPatTs_Zanini.csv",sep=",",header=TRUE,row.names=1)->freqPatTsZanini
colMeansTsZanini<-apply(freqPatTsZanini, 2 , mean, na.rm=TRUE)

#GT 10/9/16 : search apply function apply.
```

## Create overview dataframe and plot site frequency spectra 
Only synonymous, non-synomous and stop codons are considered
- for each mutation, determine whether it is synonymous, non-synonymous or creates a stop
- add information on resistance  positions

```{r}
#PSP Nov 11 2015 I removed lowerConf and upperConf here because we no longer calculate them
#x<-data.frame(num=1:984,colMeansTs,TypeOfSite,lowerConf,upperConf)
#PSP Nov 11 2015 I renamed x OverviewDFZanini and newdata OverviewDFZaniniOrderedByFreq


numsitesZanini<-length(colMeansTsZanini)
OverviewDFZanini<-data.frame(num=1:numsitesZanini,colMeansTsZanini)

OverviewDFZanini$TypeOfSite<-TypeOfSite[1:numsitesZanini]
OverviewDFZanini$TypeOfSite[1:39]<-"overlap"
OverviewDFZanini$TypeOfSite[which(consensusB != consensusC | consensusC != consensus01AE)]<-"exclude"


OverviewDFZanini$WTnt<-consensusB[1:numsitesZanini]

#Mut rates and sel coefficients
read.csv("../Data/HIVMutRates/HIVMutRates.csv")->mutrates
OverviewDFZanini$TSmutrate<-0
OverviewDFZanini$TSmutrate[OverviewDFZanini$WTnt=="a"]<-mutrates$Probability[mutrates$Nucleotide.substitution=="AG"]
OverviewDFZanini$TSmutrate[OverviewDFZanini$WTnt=="c"]<-mutrates$Probability[mutrates$Nucleotide.substitution=="CU"]
OverviewDFZanini$TSmutrate[OverviewDFZanini$WTnt=="g"]<-mutrates$Probability[mutrates$Nucleotide.substitution=="GA"]
OverviewDFZanini$TSmutrate[OverviewDFZanini$WTnt=="t"]<-mutrates$Probability[mutrates$Nucleotide.substitution=="UC"]

for (i in 1:984){
OverviewDFZanini$EstSelCoeff[i] = EstimatedS(OverviewDFZanini$TSmutrate[i],OverviewDFZanini$colMeansTsZanini[i])}

#OverviewDFZanini$EstSelCoeff= OverviewDFZanini$TSmutrate/OverviewDFZanini$colMeansTsZanini
#OverviewDFZanini$EstSelCoeff[OverviewDFZanini$EstSelCoeff>1]<-1

#WT AAs 
OverviewDFZanini$WTAA<-""
for (i in 1:numsitesZanini){
    if (i%%3==1) OverviewDFZanini$WTAA[i] = seqinr::translate(OverviewDFZanini$WTnt[c(i,i+1,i+2)])
    if (i%%3==2) OverviewDFZanini$WTAA[i] = seqinr::translate(OverviewDFZanini$WTnt[c(i-1,i,i+1)])
    if (i%%3==0) OverviewDFZanini$WTAA[i] = seqinr::translate(OverviewDFZanini$WTnt[c(i-2,i-1,i)])
}

OverviewDFZanini$MUTAA<-""
#MUT AAs
for (i in 1:numsitesZanini){
    if (i%%3==1) OverviewDFZanini$MUTAA[i] = seqinr::translate(c(transition(OverviewDFZanini$WTnt[i]),OverviewDFZanini$WTnt[c(i+1,i+2)]))
    if (i%%3==2) OverviewDFZanini$MUTAA[i] = seqinr::translate(c(OverviewDFZanini$WTnt[c(i-1)],transition(OverviewDFZanini$WTnt[i]),OverviewDFZanini$WTnt[c(i+1)]))
    if (i%%3==0) OverviewDFZanini$MUTAA[i] = seqinr::translate(c(OverviewDFZanini$WTnt[c(i-2,i-1)],transition(OverviewDFZanini$WTnt[i])))
}

write.csv(OverviewDFZanini,"../Output/OverviewSelCoeffZanini.csv")

#remove sites that are different between A and B subtypes / ALL B subtype, except pat 6 is subtype C. 
OverviewDFZanini<-OverviewDFZanini[which(consensusC==consensusB),]

#colors
OverviewDFZaniniOrderedByFreq <- OverviewDFZanini[order(OverviewDFZanini$colMeansTsZanini),] 
OverviewDFZaniniOrderedByFreq <- OverviewDFZaniniOrderedByFreq[!is.na(OverviewDFZaniniOrderedByFreq$colMeansTsZanini),] 
OverviewDFZaniniOrderedByFreq$color<-""
OverviewDFZaniniOrderedByFreq$NTcolor<-""
for (i in 1:length(OverviewDFZaniniOrderedByFreq$color)){
    if (OverviewDFZaniniOrderedByFreq$TypeOfSite[i]=="syn") OverviewDFZaniniOrderedByFreq$color[i] = "darkolivegreen3"
    if (OverviewDFZaniniOrderedByFreq$TypeOfSite[i]=="nonsyn") OverviewDFZaniniOrderedByFreq$color[i] = "red"
    if (OverviewDFZaniniOrderedByFreq$TypeOfSite[i]=="stop") OverviewDFZaniniOrderedByFreq$color[i] = "black"
    if (OverviewDFZaniniOrderedByFreq$TypeOfSite[i]=="res") OverviewDFZaniniOrderedByFreq$color[i] = "purple"
    if (OverviewDFZaniniOrderedByFreq$WTnt[i] == "a") OverviewDFZaniniOrderedByFreq$NTcolor[i] = "blue"
    if (OverviewDFZaniniOrderedByFreq$WTnt[i] == "c") OverviewDFZaniniOrderedByFreq$NTcolor[i] = "purple"
    if (OverviewDFZaniniOrderedByFreq$WTnt[i] == "g") OverviewDFZaniniOrderedByFreq$NTcolor[i] = "purple"
    if (OverviewDFZaniniOrderedByFreq$WTnt[i] == "t") OverviewDFZaniniOrderedByFreq$NTcolor[i] = "blue"
}
```


RT: Make the plots (transitions) for rankng
```{r}
#Project recreate this figure for zannini
#GT 11/7/16 added RTdata != overlap and exclude. changed pch= 16 to "|". I added legend to make the text box.
# 11/11 I think we need to change the "axis" labels to be similar to the paper's axis
#11/14 added log= "y" removed log from plot RTdata, added axis for 10^-3, removed 0.001 from RTdata change the ylim to (10^-5, 0.6)
#I tried paste("10^-3", sep="") it's the closest in giving me the the exponent label.

#function scientific()?



par(mfrow=c(1,1))
#pdf("../Output/RTRanking.pdf",width = 13, height = 10)

pdf(file="Rplot.pdf", width = 13, height = 10)
vector0.2<- array(0.2,6)
for(i in 2:6){
  vector0.2[i]= 0.1 + vector0.2[i-1]
}

vector0.02<-array(0.02, 8)
for(i in 2:8){
  vector0.02[i]= 0.01 +vector0.02[i-1]
}

vector0.002<-array(0.002, 8)
for(i in 2:8){
  vector0.002[i]= 0.001 +vector0.002[i-1]
}

vector0.0002<-array(0.0002, 8)
for(i in 2:8){
  vector0.0002[i]= 0.0001 +vector0.0002[i-1]
}

vector0.00002<-array(0.00002, 8)
for(i in 2:8){
  vector0.00002[i]= 0.00001 +vector0.00002[i-1]
}

mainvec <- c(vector0.00002, vector0.0002, vector0.002, vector0.02, vector0.2)
mainvec2 <- c(mainvec,10^(-5:-1))

RTdata<-OverviewDFZaniniOrderedByFreq
RTdata<-RTdata[RTdata$TypeOfSite!="res",]
RTdata<-RTdata[RTdata$TypeOfSite!="overlap",]
RTdata<-RTdata[RTdata$TypeOfSite!="exclude",]
plot(RTdata$colMeansTsZanini, panel.first= c(abline(h=mainvec2, lty=4, col="lightgrey")),
     main = "RT mutant frequencies",
     #ylim=c(log(0.0001),log(1)),
     ylim=c(10^-5, 0.6),
     cex=1.5, pch = "|", col=alpha(RTdata$color, .7),
     xlab = "Nucleotides ordered by mean mutation frequency",
     ylab = "Mean mutation frequency" , yaxt = "n", log="y")

axis(2, at=c(10^-5, 10^-4, 10^-3, 10^-2, 10^-1), labels = expression(10^-5, 10^-4, 10^-3, 10^-2, 10^-1),las= 1)

axis(2, at=mainvec, labels=NA,las= 1) # this makes the 8 ticks w/out the labels

legend("bottomright", c("Synonymous", "Non-synonymous", "Nonsense"), pch= "|", col = c("darkolivegreen3", "red", "black"))

dev.off()


#Use Expression function to make superscript
#grid(nx=NA, ny=NULL, col="lightgray", lwd=1, lty=6)

#dev.off()
#axis(2,labels = c(0,0.001, 0.005, 0.05, 0.1), at = log(c(0.001, 0.002, 0.006, 0.051, 0.101)),las=1)

#grid lines
#abline(v=)
#grid()

```


```{r}
#Color by NT 
plot(log(RTdata$colMeansTsZanini+0.0001), main = "RT mutant frequencies",
     ylim=c(log(0.0001),log(1)),xlim=c(0,500),cex=1.5, pch = 16, col=alpha(RTdata$NTcolor, .7), xlab = "Nucleotides ordered by mean mutation frequency", ylab = "Mean mutation frequency" , yaxt = "n")
axis(2,labels = c(0.0001, 0.001, 0.005, 0.05, 0.1,1.0), at = log(c(0.0001, 0.0011, 0.0051, 0.0501, 0.1001,1.0001)),las=1)
```
```


Show distribution of fitness effects estimated
```{r}
#GT: 10/25/16 I added col= "yellow" on abline(v=median(log10(SelcoeffA))) &(v=median(log10(SelcoeffC))) , so I can see the median line better
#This is for nonsysn log of 10 values

par(mfrow=c(2,2))
SelcoeffA<-OverviewDFZanini$EstSelCoeff[OverviewDFZanini$TypeOfSite=="nonsyn"&OverviewDFZanini$WTnt=="a"]
SelcoeffA<-SelcoeffA[!is.na(SelcoeffA)]
hist(log10(SelcoeffA),breaks= 20, xlim=c(-5,-0),ylim=c(0,30),col="red",main="A",xlab="sel coeff (10log)",ylab="count")
abline(v=median(log10(SelcoeffA)), col= "yellow") # it's about -1.365

SelcoeffC<-OverviewDFZanini$EstSelCoeff[OverviewDFZanini$TypeOfSite=="nonsyn"&OverviewDFZanini$WTnt=="c"]
SelcoeffC<-SelcoeffC[!is.na(SelcoeffC)]
hist(log10(SelcoeffC),breaks= 20, xlim=c(-5,-0),ylim=c(0,20),col="red",main="C",xlab="sel coeff (10log)",ylab="count")
abline(v=median(log10(SelcoeffC)), col= "yellow")

SelcoeffG<-OverviewDFZanini$EstSelCoeff[OverviewDFZanini$TypeOfSite=="nonsyn"&OverviewDFZanini$WTnt=="g"]
SelcoeffG<-SelcoeffG[!is.na(SelcoeffG)]
hist(log10(SelcoeffG),breaks= 20, xlim=c(-5,-0),ylim=c(0,20),col="red",main="G",xlab="sel coeff (10log)",ylab="count")
abline(v=median(log10(SelcoeffG)))

SelcoeffT<-OverviewDFZanini$EstSelCoeff[OverviewDFZanini$TypeOfSite=="nonsyn"&OverviewDFZanini$WTnt=="t"]
SelcoeffT<-SelcoeffT[!is.na(SelcoeffT)]
hist(log10(SelcoeffT),breaks= 20, xlim=c(-5,-0),ylim=c(0,20),col="red",main="T",xlab="sel coeff (10log)",ylab="count")
abline(v=median(log10(SelcoeffT)))
```

```{r}
#GT 10/25/16 I added col= "yellow" to abline(v=median(SelcoeffA)), so I can see the line better. I think might the figure for nonsynonmous on page 19.

hist(SelcoeffA,breaks= seq(0,1,by=0.01), xlim=c(0,1),ylim=c(0,50),main="A",xlab="sel coeff",ylab="count",col="red")
abline(v=median(SelcoeffA), col= "yellow")

hist(SelcoeffC,breaks= seq(0,1,by=0.01), xlim=c(0,1),ylim=c(0,10),main="C",xlab="sel coeff",ylab="count",col="green")
abline(v=median(SelcoeffC))

hist(SelcoeffG,breaks= seq(0,1,by=0.01), xlim=c(0,1),ylim=c(0,20),main="G",xlab="sel coeff",ylab="count",col="purple")
abline(v=median(SelcoeffG))


hist(SelcoeffT,breaks= seq(0,1,by=0.01), xlim=c(0,1),ylim=c(0,20),main="T",xlab="sel coeff",ylab="count",col="blue")
abline(v=median(SelcoeffT))
```



### Make an AA variability map 
```{r}
translatedB<-seqinr::translate(consensusB)
#convert the fasta files into AA sequences
translatedfasta<-c()
for (i in 1:length(listfastafiles)){ #for each fastafile 
         filename=paste("../Data/BachelerFiles/FASTAfiles/",substr(listfastafiles[i],1,6),".fasta",sep="")
	       patfasta<-read.dna(filename, format = "fasta",as.character=TRUE) #read the file       
	       translatedfasta<-rbind(translatedfasta,t(apply(patfasta,1,function(x) seqinr::translate(x))))
}
```

```{r}
#determine for each codon position, the frequency of mutations compared to subtype B
mutations<-c()
for(m in 1:length(translatedB)){
prevmut<-100-(length(which(translatedfasta[,m]==translatedB[m]))/nrow(translatedfasta)*100)
mutations<-c(mutations,prevmut)
}

# Plot for protease including resistance positions
plot(mutations[1:99],type='o',ylab='Mutation frequency (%)',xlab='AA position')
# Plot for RT including resistance positions
plot(mutations[ 100:328],type='o',ylab='Mutation frequency (%)',xlab='AA position',xaxt='n')
axis(1,(100:328)-99,label=(100:328)-99)


# Plot for RT excluding resistance positions and including resisatnce positions in red
mutationposition<-cbind(1:229,mutations[100:328])
plot(mutationposition[-c(sort(RTImuts$pos)),1],mutationposition[-c(sort(RTImuts$pos)),2],type='o',ylab='Mutation frequency (%)',xlab='AA position',xaxt='n')
points(mutationposition[c(sort(RTImuts$pos)),1],mutationposition[c(sort(RTImuts$pos)),2],col='red')
axis(1,(100:328)-99,label=(100:328)-99)
```



## Non-Syn plot 

```{r}
#GT 10/25/16 Error object 'PROdata' not found and more errors In addition: Warning message:
#In is.na(x) : is.na() applied to non-(list or vector) of type 'NULL'. 
#11/3/16 Not needed for this project Zannini.

pdf("../Output/PR-nonsyn.pdf",width = 13, height = 10)
PROdataNonSyn<-PROdata[PROdata$TypeOfSite=="nonsyn",]
#Make plot of Protease 
plot(PROdataNonSyn$num[order(PROdataNonSyn$num)],PROdataNonSyn$colMeansTs66[order(PROdataNonSyn$num)]+0.001,col=PROdataNonSyn$color[order(PROdataNonSyn$num)],pch=16,log="y",type="b", xlab = "Position in protease gene", ylab = "frequency non-synonymous transitions")#,ylim=c(0,max(PROdataNonSyn$colMeansTs66)))
#Add moving mean
for (i in 20:280){
    range = (i-20):(i+20)
#    points(i, mean(PROdataNonSyn$colMeansTs66[PROdataNonSyn$num%in%range]), col="blue",pch = 12, cex=3)
    points(i, median(PROdataNonSyn$colMeansTs66[PROdataNonSyn$num%in%range]), col="green",pch = 11, cex=2)
}
dev.off()

RTdataNonSyn<-RTdata[RTdata$TypeOfSite=="nonsyn",]
plot(RTdataNonSyn$num[order(RTdataNonSyn$num)],RTdataNonSyn$colMeansTs66[order(RTdataNonSyn$num)]+0.001,col=RTdataNonSyn$color[order(RTdataNonSyn$num)],pch=16,log="y",type="b")

#Add moving mean
for (i in 370:900){
    range = (i-20):(i+20)
   # points(i, mean(RTdataNonSyn$colMeansTs66[RTdataNonSyn$num%in%range]), col="blue",pch = 12, cex=3)
    points(i, median(RTdataNonSyn$colMeansTs66[RTdataNonSyn$num%in%range]), col="green",pch = 11, cex=3)
}

i=370
    range = (i-20):(i+20)
```


## Syn plot 

```{r}
pdf("../Output/PR-syn.pdf",width = 13, height = 10)
PROdataSyn<-PROdata[PROdata$TypeOfSite=="syn",]
#Make plot of Protease 
plot(PROdataSyn$num[order(PROdataSyn$num)],PROdataSyn$colMeansTs66[order(PROdataSyn$num)]+0.001,col=PROdataSyn$color[order(PROdataSyn$num)],pch=16,
     log="y",
     type="b")#,ylim=c(0,max(PROdataNonSyn$colMeansTs66)))

#Add moving mean
for (i in 20:280){
    range = (i-10):(i+10)
    points(i, median(PROdataSyn$colMeansTs66[PROdataSyn$num%in%range]), col="blue",pch = 11, cex=2)
}
dev.off()
```

```{r}
#GT: 10/26/16 Error in xy.coords(x,y, xlabel, log): 'x' and 'y' length differ, looked up what order function does

RTdataNonSyn<-RTdata[RTdata$TypeOfSite=="nonsyn",]
plot(RTdataNonSyn$num[order(RTdataNonSyn$num)],RTdataNonSyn$colMeansTs66[order(RTdataNonSyn$num)]+0.001,col=RTdataNonSyn$color[order(RTdataNonSyn$num)],pch=16,log="y",type="b")

#Add moving mean
for (i in 370:900){
    range = (i-20):(i+20)
   # points(i, mean(RTdataNonSyn$colMeansTs66[RTdataNonSyn$num%in%range]), col="blue",pch = 12, cex=3)
    points(i, median(RTdataNonSyn$colMeansTs66[RTdataNonSyn$num%in%range]), col="green",pch = 11, cex=3)
}

i=370
    range = (i-20):(i+20)
```

```{r}
#GT 10/26/16 still receiving error messages about xy.coord(x,y): different length In is.na(x) : is.na() applied to non-(list or vector) of type 'NULL'

pdf("../Output/RT-syn.pdf",width = 13, height = 10)
RTdataSyn<-RTdata[RTdata$TypeOfSite=="syn",]
#Make plot of RTtease 
plot(RTdataSyn$num[order(RTdataSyn$num)],RTdataSyn$colMeansTs66[order(RTdataSyn$num)]+0.001,col=RTdataSyn$color[order(RTdataSyn$num)],pch=16,
     log="y",
     type="b", 
    ylim = c(0.001,0.06)    )#,ylim=c(0,max(RTdataNonSyn$colMeansTs66)))
#Add moving mean
for (i in 370:900){
    range = (i-10):(i+10)
    points(i, median(RTdataSyn$colMeansTs66[RTdataSyn$num%in%range])+0.001, col="blue",pch = 11, cex=2)
}
dev.off()
```

```{r}
AlldataSyn<-rbind(PROdata[PROdata$TypeOfSite=="syn",],RTdata[RTdata$TypeOfSite=="syn",])
plot(AlldataSyn$num[order(AlldataSyn$num)],AlldataSyn$colMeansTs66[order(AlldataSyn$num)]+0.001,col=AlldataSyn$color[order(AlldataSyn$num)],pch=16,
     log="y",
     type="b", 
    ylim = c(0.001,0.06)    )#,ylim=c(0,max(RTdataNonSyn$colMeansTs66)))
#Add line for AA 82-107 from Mayrose paper
lines(c(82*3,107*3),c(0.002,0.002),col=2,lwd=2)
lines(c(7*3,31*3),c(0.002,0.002),col=2,lwd=2)


```

## RT
RTdataNonSyn<-RTdata[RTdata$TypeOfSite=="nonsyn",]
plot(RTdataNonSyn$num[order(RTdataNonSyn$num)],RTdataNonSyn$colMeansTs66[order(RTdataNonSyn$num)]+0.001,col=RTdataNonSyn$color[order(RTdataNonSyn$num)],pch=16,log="y",type="b")

#Add moving mean
for (i in 370:900){
    range = (i-20):(i+20)
    points(i, mean(RTdataNonSyn$colMeansTs66[RTdataNonSyn$num%in%range]), col="blue",pch = 12, cex=3)
    points(i, median(RTdataNonSyn$colMeansTs66[RTdataNonSyn$num%in%range]), col="green",pch = 11, cex=3)
}

i=370
    range = (i-20):(i+20)
    
```    



# Next steps 
Going to randomisations
- frequency-ordered figure: does the coloring scheme hold?
  
- position-ordered figure: does the pattern hold?  
    
###For protease, which are the most deleterious non-syn sites? 
To be done 
Interesting? why? for therapy or vaccine. 

###For protease, which are the most neutral non-syn sites? 
corresponding to the expected frequency based on sequence length * mutation rate? Frequency should flucatuate aroudn the expected value? 

#For protease, which are the most deleterious syn sites? 
To be done : green colored sites with low frequ
Interesting why?  HIV replication?





